---
title: "p8105_hw6_yg2964"
output: github_document
date: "2024-12-02"
---

```{r}
library(tidyverse)
library(modelr)
library(ggplot2)
library(mgcv)
library(broom)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(127)
```

### Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

get_resample = function(df) {
  sample_frac(df, replace = TRUE)
}

extract_metrics = function(model) {
  intercept = tidy(model) |> filter(term == "(Intercept)") |> pull(estimate)
  slope = tidy(model) |> filter(term == "tmin") |> pull(estimate)
  
  tibble(
    rsquared = glance(model)$r.squared,
    coef_product = log(abs(intercept * slope))
  )
}

resampling_sets = 
  tibble(rep = 1:5000) |>
  mutate(
    resamples = map(rep, ~get_resample(weather_df))
  )

resample_results = 
  resampling_sets |>
  mutate(
    fits = map(resamples, ~lm(tmax ~ tmin, data = .)),
    metrics = map(fits, extract_metrics)
  ) |>
  select(-resamples, -fits) |>
  unnest(metrics)

# Get confidence bounds
conf_intervals = resample_results |>
  summarize(across(everything(), 
                  list(lower = ~quantile(., 0.025),
                       upper = ~quantile(., 0.975))))

# Create visualizations
viz1 = resample_results |>
  ggplot(aes(x = rsquared)) +
  geom_histogram(fill = "seagreen3", color = "seagreen4", alpha = 0.8) +
  labs(title = "R² Distribution", 
       x = "R-squared Value", 
       y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

viz2 = resample_results |>
  ggplot(aes(x = coef_product)) +
  geom_histogram(fill = "orchid3", color = "orchid4", alpha = 0.8) +
  labs(title = "log(β₀*β₁) Distribution", 
       x = "Log Product of Coefficients", 
       y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

viz1
viz2
```

R² Distribution:
Centered around 0.91 with values ranging from 0.88 to 0.94, showing a compact and symmetric distribution that indicates a strong and stable linear relationship between tmax and tmin.
log(β₀*β₁) Distribution:
Centered around 2.02 with values ranging from 1.95 to 2.10, displaying a symmetric distribution that confirms stable, positive coefficient products and supports the positive correlation between variables.